---
title: "Priors, P-values, and Pfizer: Reanalyzing the BNT162b2 Vaccine Efficacy Trial"
author: "Joe Clair, Jay Kumar, and Andrew Tang"
date: "March 10, 2026"
output:
  html_document:
    df_print: paged
urlcolor: blue
header-includes:
- \usepackage{amsmath,amsfonts,amssymb}
- \usepackage{setspace} \doublespacing
fontsize: 11pt
---

```{r setup, include=FALSE}
#Use this code chunk to include libraries, and set global options.
```

# Abstract
Write your abstract here.

# Keywords
*Vaccine Efficacy*, *COVID-19*, *Prior Elicitation*, *Bayesian Analysis*

\newpage

# Introduction / Background

The COVID-19 pandemic, caused by SARS-CoV-2, was one of the most disruptive global health crises 
in modern history. The pandemic impacted global health substantially by not only claiming an estimated 7.1 million 
lives worldwide (WHO, 2022), but also triggering widespread economic disruption and social issues, 
including school closures and lockdowns across the globe.

In response, Pfizer and BioNTech developed BNT162b2, a two-dose mRNA vaccine, which was a new approach 
compared to traditional vaccine platforms. The vaccine received FDA Emergency Use Authorization in 
December 2020 after a large-scale randomized, double-blind, placebo-controlled Phase 3 trial 
(Polack et al., 2020). The trial was designed to run until 170 confirmed COVID-19 cases were observed, 
ultimately enrolling 34,922 participants split equally between vaccine and placebo groups. Vaccine 
efficacy $\psi$, defined as the proportional reduction in infection risk relative to placebo, must 
exceed 30% to meet FDA minimum approval standards. The results are summarized in Table 1 below.

#Add table here? It says so in the instructions

To evaluate vaccine efficacy, each company who developed a vaccine employed statistical methods to analyze their data. 
Pfizer used a Bayesian beta-binomial model and reported 95% efficacy (Polack et al., 2020), while another company,
Moderna, modeled their data using a frequentist approach on a similar mRNA vaccine and reported 
94.1% efficacy (Baden et al., 2021). The fact that two different statistical frameworks applied to 
similar mRNA vaccines had similar conclusions was promising, suggesting the high efficacy 
of this mRNA holds strong no matter the statistical framework. However, Pfizer's Bayesian analysis did 
not go without scrutiny. Senn (2021) argues that their prior distribution was
reverse-engineered from the FDA's 30% threshold rather than reflecting genuine pre-trial belief. 
We therefore independently reanalyze the Pfizer trial data using both Bayesian and frequentist 
methods, this time using a prior that is more in line with the Bayesian framework of eliciting our prior beliefs. After our test is done, we will see if the same conclusions hold. Specifically, we test:

$$H_0: \psi = 0.30 \quad \text{vs} \quad H_1: \psi > 0.30$$

```{r label="important_R_code", eval=TRUE, echo=FALSE}
# You can reference your code in the appendix (sample here).
```

# Statistical Methods

## Model
Suppose we define the random variable T as the number in the Vaccine group from the n = 170 Covid 19 cases

\begin{center}
    \begin{tabular}{|c|c|c|}
        \hline
        Group & Cases & Sample Size \\
        \hline
        BNT162b2 & 2 & 17411 \\
        \hline
        Placebo & 168 & 17511 \\
        \hline
        Total & 170 & 34922 \\
        \hline
    \end{tabular}
\end{center}

The distribution of T can be well approximated by a binomial distribution which is proved in the appendix

Then,

$$T \sim Binom(n = 170, \pi) $$
Where $\pi$ = P(vaccine | Covid 19 Case) 

## Likelihood Inference
Detail the likelihood approach.

## Bayesian Inference

Before seeing the data, we set two prior beliefs about vaccine efficacy $\psi$. We chose 
$P(\psi \geq 0.3) = 0.10$ because mRNA vaccines had never been approved before, and most 
Phase 3 vaccine trials fail, so we were skeptical the vaccine would clear the FDA's 30% bar. 
We chose $P(\psi \geq 0) = 0.60$ given promising immune responses observed in Phase 1/2 trials (Polack et al., 2020). Following Senn (2021), who argues Pfizer's prior was reverse-engineered from the FDA 
threshold rather than reflecting real prior beliefs, we created our beliefs directly 
rather than working backwards from the approval standard. We now will use the `beta.select` function in R in order
to generate a beta distribution from our new prior beliefs.

Under our prior beliefs,
\begin{align*}
P (\psi \geq 0.3) = 0.1, \ \ \ P (\psi \geq 0) = 0.6.
\end{align*}

When $\psi = 0.3$,
\begin{align*}
\pi = \frac{1-0.3}{2-0.3} = \frac{0.7}{1.7} = \frac{7}{17}.
\end{align*}
Thus,
\begin{align*}
P(\psi \geq 0.3) = 0.1
\quad \Longleftrightarrow \quad
P\left(\pi \leq \frac{7}{17}\right) = 0.1.
\end{align*}

When $\psi = 0$,
\begin{align*}
\pi = \frac{1-0}{2-0} = \frac{1}{2}.
\end{align*}
Thus,
\begin{align*}
P(\psi \geq 0) = 0.6
\quad \Longleftrightarrow \quad
P\left(\pi \leq \frac{1}{2}\right) = 0.6.
\end{align*}

So, our conditions for $\pi$ are
\begin{align*}
P\left(\pi \leq \frac{7}{17}\right) = 0.1,
\qquad
P\left(\pi \leq \frac{1}{2}\right) = 0.6.
\end{align*}

Using this to create a beta distribution gives $Beta (36.40, 38.58)$ as the prior.
```{r}
library(LearnBayes)

beta.select(
  quantile1 = list(p = 0.10, x = 7/17),
  quantile2 = list(p = 0.60, x = 0.50)
)
```
Because the Beta distribution is conjugate to the Binomial, the posterior distribution of $\pi$ is also a Beta distribution. 
\[
T \mid \pi \sim \text{Binomial}(n,\pi)
\]
and
\[
\pi \sim \text{Beta}(a,b),
\]
then the posterior distribution is
\[
\pi \mid T=t \sim \text{Beta}(a+t,\; b+n-t).
\]
For our data, $n=170$ and $t=8$, so the posterior distribution is
\[
\pi \mid T=8 \sim \text{Beta}(a+8,\; b+162).
\]


# Results
Present your findings. 

# Discussion / Conclusion
Discuss / conclude here.

# Bibliography

Baden, L. R., El Sahly, H. M., Essink, B., Kotloff, K., Frey, S., Novak, R., ... & Zaks, T. (2021). Efficacy and safety of the mRNA-1273 SARS-CoV-2 vaccine. *New England Journal of Medicine*, 384(5), 403--416.

Polack, F. P., Thomas, S. J., Kitchin, N., Absalon, J., Gurtman, A., Lockhart, S., ... & Jansen, K. U. (2020). Safety and efficacy of the BNT162b2 mRNA Covid-19 vaccine. *New England Journal of Medicine*, 383(27), 2603--2615.

Senn, S. (2021). *Beta testing: The Pfizer/BioNTech statistical analysis of their Covid-19 vaccine trial*. Error Statistics Philosophy. Retrieved from https://errorstatistics.com/2021/01/17/s-senn-beta-testing-the-pfizer-biontech-statistical-analysis-of-their-covid-19-vaccine-trial-guest-post/

World Health Organization. (2022). *COVID-19 dashboard*. Retrieved from https://data.who.int/dashboards/covid19/deaths?n=c

Two things to note:

# Appendix

## Code
```{r ref.label = "important_R_code", eval=FALSE}
```

## Proofs
If applicable, include detailed mathematical derivations or additional theoretical explanations.