---
title: "Changed"
author: "Joe Clair, Jay Kumar, and Andrew Tang"
date: Date
output: pdf_document
urlcolor: blue
header-includes:
- \usepackage{amsmath,amsfonts,amssymb}
- \usepackage{setspace} \doublespacing
fontsize: 11pt
---

```{r setup, include=FALSE}
#Use this code chunk to include libraries, and set global options.
```

# Abstract
Write your abstract here.

# Keywords
*Keyword 1*, *Keyword 2*, *Keyword 3*, *Keyword 4*

\newpage

# Introduction / Background
Provide an introduction with background information.

```{r label="important_R_code", eval=TRUE, echo=FALSE}
# You can reference your code in the appendix (sample here).
```

# Statistical Methods

## Model
Suppose we define the random variable T as the number in the Vaccine group from the n = 170 Covid 19 cases

\begin{center}
    \begin{tabular}{|c|c|c|}
        \hline
        Group & Cases & Sample Size \\
        \hline
        BNT162b2 & 2 & 17411 \\
        \hline
        Placebo & 168 & 17511 \\
        \hline
        Total & 170 & 34922 \\
        \hline
    \end{tabular}
\end{center}

The distribution of T can be well approximated by a binomial distribution which is proved in the appendix

Then,

$$T \sim Binom(n = 170, \pi) $$
Where $\pi$ = P(vaccine | Covid 19 Case) 

## Likelihood Inference
Detail the likelihood approach.

## Bayesian Inference
Detail the Bayesian approach.

# Results
Present your findings. 

# Discussion / Conclusion
Discuss / conclude here.

# Bibliography

Brown, B. (2024). *Lecture Title*. Lecture slides, Course Name, University Name.

Doe, J. (2020). Title of the Paper. *Journal Name*, 12(3), 45-67.

Last, F., & Last, F. (2025). *Book Title*. Publisher.

Smith, A., & Johnson, C. (2023). *Title of the Online Article*. Retrieved from https://www.example.com.

# Appendix

## Code
```{r ref.label = "important_R_code", eval=FALSE}
```

## Proofs
If applicable, include detailed mathematical derivations or additional theoretical explanations.