---
title: "Priors, P-values, and Pfizer: Reanalyzing the BNT162b2 Vaccine Efficacy Trial"
author: "Joe Clair, Jay Kumar, and Andrew Tang"
date: "March 10, 2026"
output:
  pdf_document: default
  html_document:
    df_print: paged
urlcolor: blue
header-includes:
- \usepackage{amsmath,amsfonts,amssymb}
- \usepackage{setspace} \doublespacing
fontsize: 11pt
---

```{r setup, include=FALSE}
#Use this code chunk to include libraries, and set global options.
library(fastR2)
library(ggplot2)
library(HDInterval)
```

# Abstract
The COVID-19 pandemic led to rapid vaccine development, especially in Pfizer-BioNTech's 
BNT162b2 mRNA vaccine receiving FDA Emergency Use Authorization in December 2020. We 
independently reanalyze their Phase 3 trial data using both Bayesian and Frequentist methods 
to test whether vaccine efficacy, $\psi$, exceeds the FDA minimum threshold of 30\%. This addresses
concerns raised by Senn (2021) that Pfizer's original prior was reverse-engineered from this 
threshold rather than reflecting genuine pre-trial belief. Under both approaches, the data 
provided overwhelming evidence that vaccine efficacy exceeds 30\%, with a frequentist MLE of 
$\hat{\psi} = 0.951$ and a Bayesian posterior median of $0.780$. Despite using a more 
skeptical prior than Pfizer, our conclusions are consistent with their reported 95\% efficacy, 
suggesting the results are not only strong with both frameworks, but even a more relevant prior.

# Keywords
*Vaccine Efficacy*, *COVID-19*, *Prior Elicitation*, *Bayesian Analysis*

\newpage

# Introduction / Background

The COVID-19 pandemic, caused by SARS-CoV-2, was one of the most disruptive global health crises 
in modern history. The pandemic impacted global health substantially by not only claiming an estimated 7.1 million 
lives worldwide (WHO, 2022), but also triggering widespread economic disruption and social issues, 
including school closures and lockdowns across the globe.

In response, Pfizer and BioNTech developed BNT162b2, a two-dose mRNA vaccine, which was a new approach 
compared to traditional vaccine platforms. The vaccine received FDA Emergency Use Authorization in 
December 2020 after a large-scale randomized, double-blind, placebo-controlled Phase 3 trial 
(Polack et al., 2020). The trial was designed to run until 170 confirmed COVID-19 cases were observed, 
ultimately enrolling 34,922 participants split equally between vaccine and placebo groups. Vaccine 
efficacy $\psi$, defined as the proportional reduction in infection risk relative to placebo, must 
exceed 30% to meet FDA minimum approval standards. The results are summarized in Table 1 below.

\begin{center}
    \begin{tabular}{|c|c|c|}
        \hline
        Group & Cases & Sample Size \\
        \hline
        BNT162b2 & 8 & 17411 \\
        \hline
        Placebo & 162 & 17511 \\
        \hline
        Total & 170 & 34922 \\
        \hline
    \end{tabular}
\end{center}

To evaluate vaccine efficacy, each company who developed a vaccine employed statistical methods to analyze their data. 
Pfizer used a Bayesian beta-binomial model and reported 95% efficacy (Polack et al., 2020), while another company,
Moderna, modeled their data using a frequentist approach on a similar mRNA vaccine and reported 
94.1% efficacy (Baden et al., 2021). The fact that two different statistical frameworks applied to 
similar mRNA vaccines had similar conclusions was promising, suggesting the high efficacy 
of this mRNA holds strong no matter the statistical framework. However, Pfizer's Bayesian analysis did 
not go without scrutiny. Senn (2021) argues that their prior distribution was
reverse-engineered from the FDA's 30% threshold rather than reflecting genuine pre-trial belief. 
We therefore independently reanalyze the Pfizer trial data using both Bayesian and frequentist 
methods, this time using a prior that is more in line with the Bayesian framework of eliciting our prior beliefs. After our test is done, we will see if the same conclusions hold. Specifically, we test:

$$H_0: \psi = 0.30 \quad \text{vs} \quad H_1: \psi > 0.30$$


# Statistical Methods

## Model

The parameter of interest for this study is,

$$1 - \frac{ \hat \pi_v}{ \hat \pi_p}$$
This gives us the efficacy of the vaccine against Covid-19, where $\hat \pi_v$ is the probability of Covid-19 with the vaccine and $\hat \pi_p$ is the probability of Covid-19 with the placebo.

Suppose we define the random variable T as the number in the Vaccine group from the n = 170 Covid-19 cases

The distribution of T can be estimated by a binomial distribution which is proved in the appendix

Then,

$$T \sim Binom(n = 170, \pi) $$
Where $\pi$ = P(vaccine | Covid 19 Case) which is equal to,

$$ \pi = \frac{n_1\pi_v}{n_1\pi_v + n_2\pi_p}$$
Since $n_1 \approx n_2$,  $\pi$ becomes,
$$\pi  = \frac{\pi_v}{\pi_v + \pi_p}$$
Our parameter of interest is the efficacy which as defined is,
$$ \psi = 1 - \frac{ \pi_v}{  \pi_p} = \frac{\pi_v -\pi_p }{\pi_p}$$
When can represent this in terms of our binomial probability $\pi$ as,

$$\psi = \frac{1 - 2\pi}{1 - \pi}$$
Since the FDA requires at least a 30% efficacy for the vaccine to be approved our hypothesis that we will test will be,

$H_o : \psi = 0.3$ versus $H_1 : \psi > 0.3$

## Likelihood Inference
We will calculate the parameter $\psi$ using the likelihood estimator and since $T \sim Binom(n = 170, \pi)$ the Likelihood function in terms of $\pi$ will be,

$$L(\pi) = \binom{n}{k}\pi^k(1-\pi)^{n - k} =\binom{170}{8}\pi^8(1-\pi)^{162} $$
We solve this to find the MLE of $\pi$ and since MLEs are invariant to transformation we can solve for $\psi$.

We can represent $\psi$ with $\pi$ as the following,
$$\pi = \frac{1 - \psi}{2 - \psi}$$



To calculate the confidence interval for our estimate, we can use the theorem that our estimate is normally distributed with $\psi_0$ as the mean and the standard error as $\frac{1}{\sqrt{I(\psi_0)}}$, where $I(\psi_0)$ is the expected fisher information.
To show that n is large enough and that regularity condition holds for this theorem to be used we need to verify that the log-likelihood function can be well approximated by a second order Taylor series around the MLE. Since $\psi$ is a transformation of $\pi$, we verify the quadratic 
approximation using the log-likelihood in terms of $\pi$, which is equivalent  by the invariance property of the MLE.

```{r, echo=FALSE, warning=FALSE}
loglik.pi <- function(pi, v, n) {
  ifelse((pi <= 0 | pi >= 1), NA,
    v * log(pi) + (n - v) * log(1 - pi)
  )
}

ml.pi <- maxLik2(loglik = loglik.pi,
                 start = 0.5,
                 v = 8,
                 n = 170)

plot(ml.pi) +
  labs(title = "Second Order Approximation of Log-Likelihood",
       subtitle = "n = 170, t = 8",
       x = expression(pi),
       y = "Log-Likelihood")
```

This graph shows that the log-Likelihood function can be well approximated by a second order Taylor series around the MLE so the regularity conditions hold so we justify our use of the normal distribution of the MLE to calculate the estimate. 

The fisher expected fisher information is given by,

$$I(\psi_0) = E[-\frac{d^2}{d\psi^2}l^*(\psi)]_{\psi = \psi_0}$$
We can calculate a 95% confidence interval with,
$$\hat \psi^{mle} \pm z_{\alpha/2}\sqrt{\frac{1}{I(\hat \psi^{mle})}}$$

We will also calculate the interval through a parametric bootstrap percentile interval.

To do this we will follow the steps of first repeatedly creating new samples from the distribution,
$$T \sim Binom(n = 170,  \hat \pi)$$
We re-sample from the binomial distribution using our estimated $\pi$ from the original sample of the trials. Then with each sample we will calculate our estimate $\psi$ and then for the 95% confidence interval will take the 2.5th and 97.5th percentile to create our confidence interval.

To assess whether the data provide sufficient evidence against $H_0$, we use the likelihood ratio test statistic,

$$W = 2[\ell^*(\hat{\psi}^{mle}) - \ell^*(\psi_0)]$$
which measures how much more likely the observed data are under the MLE compared to the null 
value $\psi_0 = 0.3$. Large values of $W$ indicate the null hypothesis is not likely given 
the data. Under $H_0$, $W$ follows an approximate $\chi^2_1$ distribution for large $n$, 
which we use to compute a P-value. We also compute an empirical P-value by simulating the 
distribution of $W$ under $H_0$ directly.

## Bayesian Inference

Before seeing the data, we set two prior beliefs about vaccine efficacy $\psi$. We chose 
$P(\psi \geq 0.3) = 0.10$ because mRNA vaccines had never been approved before, and most 
Phase 3 vaccine trials fail, so we were skeptical the vaccine would clear the FDA's 30% bar. 
We chose $P(\psi \geq 0) = 0.60$ given promising immune responses observed in Phase 1/2 trials (Polack et al., 2020). Following Senn (2021), who argues Pfizer's prior was reverse-engineered from the FDA 
threshold rather than reflecting real prior beliefs, we created our beliefs directly 
rather than working backwards from the approval standard. We now will use the `beta.select` function in R in order
to generate a beta distribution from our new prior beliefs.

Under our prior beliefs,
\begin{align*}
P (\psi \geq 0.3) = 0.1, \ \ \ P (\psi \geq 0) = 0.6.
\end{align*}

When $\psi = 0.3$,
\begin{align*}
\pi = \frac{1-0.3}{2-0.3} = \frac{0.7}{1.7} = \frac{7}{17}.
\end{align*}
Thus,
\begin{align*}
P(\psi \geq 0.3) = 0.1
\quad \Longleftrightarrow \quad
P\left(\pi \leq \frac{7}{17}\right) = 0.1.
\end{align*}

When $\psi = 0$,
\begin{align*}
\pi = \frac{1-0}{2-0} = \frac{1}{2}.
\end{align*}
Thus,
\begin{align*}
P(\psi \geq 0) = 0.6
\quad \Longleftrightarrow \quad
P\left(\pi \leq \frac{1}{2}\right) = 0.6.
\end{align*}

So, our conditions for $\pi$ are
\begin{align*}
P\left(\pi \leq \frac{7}{17}\right) = 0.1,
\qquad
P\left(\pi \leq \frac{1}{2}\right) = 0.6.
\end{align*}

Using this to create a beta distribution gives $Beta (36.40, 38.58)$ as the prior.
```{r}
library(LearnBayes)

beta.select(
  quantile1 = list(p = 0.10, x = 7/17),
  quantile2 = list(p = 0.60, x = 0.50)
)
```
Because the Beta distribution is conjugate to the Binomial, the posterior distribution of $\pi$ is also a Beta distribution. 
\[
T \mid \pi \sim \text{Binomial}(n,\pi)
\]
and
\[
\pi \sim \text{Beta}(a,b),
\]
then the posterior distribution is
\[
\pi \mid T=t \sim \text{Beta}(a+t,\; b+n-t).
\]
For our data, $n=170$ and $t=8$, so the posterior distribution is
\[
\pi \mid T=8 \sim \text{Beta}(a+8,\; b+162).
\]

```{r, echo = FALSE}
x <- 8
n <- 170
a <- 36.40
b <- 38.58

ggplot() +
  geom_function(
    fun = dbeta,
    mapping = aes(color = "prior"),
    args = list(shape1 = a, shape2 = b),
    xlim = c(0, 1)
  ) +
  geom_function(
    fun = dbeta,
    mapping = aes(color = "posterior"),
    args = list(shape1 = a + x, shape2 = b + n - x),
    xlim = c(0, 1)
  ) +
  scale_color_manual(name = "Distribution", values = c("prior" = "blue", "posterior" = "red")) +
  labs(
    title = "Pfizer Bayesian Prior and Posterior",
    subtitle = "Observed x = 8 out of n = 170",
    x = expression(pi),
    y = "Density"
  )
```

The Bayesian posterior summary for $\pi$ is then:

\begin{table}[h!]
\centering
\begin{tabular}{|l|c|}
\hline
\textbf{Quantity} & \textbf{Result} \\
\hline
Posterior Median & 0.180371 \\
\hline
95\% Equal-Tail & $(0.1356515, \ 0.2317538)$ \\
\hline
95\% HPDI & $(0.1340409, \ 0.2299012)$ \\
\hline
\end{tabular}
\end{table}

To convert the posterior summaries from $\pi$ to $\psi$, we use
\[
\psi = \frac{1-2\pi}{1-\pi}.
\]

Using the posterior median of $\pi$, we have
\[
\psi_{\text{median}}
= \frac{1 - 2(0.180371)}{1 - 0.180371}
= \frac{1 - 0.360742}{0.819629}
= \frac{0.639258}{0.819629}
= 0.779936.
\]

For the 95\% equal-tail interval, we transform each endpoint:
\[
\psi_{\text{lower}}
= \frac{1 - 2(0.2317538)}{1 - 0.2317538}
= \frac{1 - 0.4635076}{0.7682462}
= \frac{0.5364924}{0.7682462}
= 0.698334,
\]
and
\[
\psi_{\text{upper}}
= \frac{1 - 2(0.1356515)}{1 - 0.1356515}
= \frac{1 - 0.2713030}{0.8643485}
= \frac{0.7286970}{0.8643485}
= 0.843059.
\]

Similarly, transforming the HPDI endpoints gives
\[
\psi_{\text{lower}}
= \frac{1 - 2(0.2299012)}{1 - 0.2299012}
= 0.701465,
\qquad
\psi_{\text{upper}}
= \frac{1 - 2(0.1340409)}{1 - 0.1340409}
= 0.845211.
\]

The Bayesian posterior summary for $\psi$ is then:

\begin{table}[h!]
\centering
\begin{tabular}{|l|c|}
\hline
\textbf{Quantity} & \textbf{Result} \\
\hline
Posterior Median & 0.779936 \\
\hline
95\% Equal-Tail & $(0.698334,\ 0.843059)$ \\
\hline
95\% HPDI & $(0.701465,\ 0.845211)$ \\
\hline
\end{tabular}
\end{table}

To evaluate the Bayesian evidence against the efficacy benchmark of 30\%, we calculate
\[
P(\psi \leq 0.3 \mid t).
\]
Using the relationship
\[
\pi = \frac{1-\psi}{2-\psi},
\]
we first convert the statement \(\psi \leq 0.3\) into a statement about \(\pi\). 
\begin{align*}
\psi &\leq 0.3, \\
\frac{1-2\pi}{1-\pi} &\leq 0.3, \\
1 - 2\pi &\leq 0.3(1-\pi), \\
1 - 2\pi &\leq 0.3 - 0.3\pi, \\
0.7 - 2\pi &\leq -0.3\pi, \\
0.7 &\leq 1.7\pi, \\
\pi &\geq \frac{0.7}{1.7}.
\end{align*}

So,
\[
P(\psi \leq 0.3 \mid t) = P\left(\pi \geq \frac{0.7}{1.7} \mid t\right).
\]


# Results

### Frequentist
We will now discuss the results of our inference using Likelihood.
Using the MLE we calculate our parameter of interest in the appendix as,
$$\hat \psi^{mle} =  \frac{154}{162} = 0.95061$$
We use the normal distribution of our to estimate the 95% Confidence interval,
which the calculation in the appendix estimates as,
(0.9171746, 0.9840599)

The parametric bootstrap estimates the interval as,
(0.9102564, 0.9820359 )


Both hypothesis tests provided overwhelming evidence against $H_0: \psi = 0.30$. The 
likelihood ratio test statistic was $W = 121.6012$, which under the 
$\chi^2_1$ approximation yields a P-value essentially equal to zero. The empirical P-value, 
obtained by simulating 10,000 samples from the null distribution of $T \sim Binom(170, \pi_0)$ 
where $\pi_0 = 7/17$, was similarly $p_{emp} = 0$. Both results provide evidence that the true vaccine efficacy exceeds the FDA minimum threshold 
of 30%.

### Bayesian

The posterior probability $P \left( \psi \leq 0.3 \mid t \right)$ corresponds to $P \left(\pi \geq \frac{0.7}{1.7} \mid t \right)$ which is $4.367864e-15$. This posterior probability is essentially 0, indicating that given the observed data, there is convincing evidence that vaccine efficacy is above $0.3$.

This conclusion is also supported by the posterior median and the 95\% credible intervals for \(\psi\). The posterior median is \(0.779936\), and both the 95\% equal-tail interval \((0.698334,\ 0.843059)\) and the 95\% HPDI \((0.701465,\ 0.845211)\) lie entirely above \(0.3\). 

# Discussion / Conclusion

Both the frequentist and Bayesian analyses provide overwhelming evidence that the BNT162b2 
vaccine efficacy exceeds the FDA minimum threshold of 30\%. The Frequentist MLE of 
$\hat{\psi} = 0.951$ and the Bayesian posterior median of $0.780$ both point to a highly 
effective vaccine, and all confidence and credible intervals lie entirely above 30\%.
The two methods differ in their point estimates, which is expected given our deliberately 
skeptical prior. Our Bayesian analysis assigned only a 10\% prior probability that the vaccine 
would meet the FDA threshold, yet the posterior was still overwhelmingly in favor of high 
efficacy.

This allows us to note something else Senn (2021) mentioned in his blog critiquing the Pfizer prior. While we agree that Pfizer's 
Beta(0.700102, 1) prior was not reflective of genuine pre-trial belief, our reanalysis shows 
that the choice of prior had little practical impact on the conclusions. The trial data was so decisive that both a potentially lenient prior 
and a skeptical one end up with essentially the same result. Pfizer's reported 95\% efficacy 
with credible interval $[90.3\%, 97.6\%]$ is consistent with our findings under both 
frameworks, suggesting the high efficacy of BNT162b2 is a strong conclusion regardless of 
the statistical approach and prior taken.

# Bibliography

Baden, L. R., El Sahly, H. M., Essink, B., Kotloff, K., Frey, S., Novak, R., ... & Zaks, T. (2021). Efficacy and safety of the mRNA-1273 SARS-CoV-2 vaccine. *New England Journal of Medicine*, 384(5), 403--416.

Polack, F. P., Thomas, S. J., Kitchin, N., Absalon, J., Gurtman, A., Lockhart, S., ... & Jansen, K. U. (2020). Safety and efficacy of the BNT162b2 mRNA Covid-19 vaccine. *New England Journal of Medicine*, 383(27), 2603--2615.

Senn, S. (2021). *Beta testing: The Pfizer/BioNTech statistical analysis of their Covid-19 vaccine trial*. Error Statistics Philosophy. Retrieved from https://errorstatistics.com/2021/01/17/s-senn-beta-testing-the-pfizer-biontech-statistical-analysis-of-their-covid-19-vaccine-trial-guest-post/

World Health Organization. (2022). *COVID-19 dashboard*. Retrieved from https://data.who.int/dashboards/covid19/deaths?n=c

# Appendix

## Code
```{r ref.label = "important_R_code", eval=FALSE}
```

```{r}
#Calculating W and then Chi Sq approximation P-value
n <- 170
t <- 8
psi_hat <- (n - 2*t) / (n - t)
psi_0 <- 0.3

loglik.psi <- function(psi, t, n) {
  t * log(1 - psi) - n * log(2 - psi)
}

W_obs <- 2 * (loglik.psi(psi_hat, t, n) - loglik.psi(psi_0, t, n))
W_obs

1 - pchisq(W_obs, df = 1)
```

```{r, cache=TRUE}
#Empirical P-value calcuation
set.seed(342)
B <- 10000
n <- 170
psi_0 <- 0.3
pi_0 <- (1 - psi_0) / (2 - psi_0)  # null value of pi

w_star <- replicate(B, {
  t_star <- rbinom(1, size = n, prob = pi_0)
  psi_hat_star <- (n - 2*t_star) / (n - t_star)
  2 * (loglik.psi(psi_hat_star, t_star, n) - loglik.psi(psi_0, t_star, n))
})

p_emp <- mean(w_star >= W_obs)
p_emp

df <- data.frame(w_star = w_star)
ggplot(df, aes(x = w_star)) +
  geom_histogram() +
  geom_vline(xintercept = W_obs) +
  labs(
    title = "Empirical Null Distribution of W*",
    x = "W*",
    y = "Frequency"
  )
```

```{r}
#Prior and Posterior graph
x <- 8
n <- 170
a <- 36.40
b <- 38.58

ggplot() +
  geom_function(
    fun = dbeta,
    mapping = aes(color = "prior"),
    args = list(shape1 = a, shape2 = b),
    xlim = c(0, 1)
  ) +
  geom_function(
    fun = dbeta,
    mapping = aes(color = "posterior"),
    args = list(shape1 = a + x, shape2 = b + n - x),
    xlim = c(0, 1)
  ) +
  scale_color_manual(name = "Distribution", values = c("prior" = "blue", "posterior" = "red")) +
  labs(
    title = "Pfizer Bayesian Prior and Posterior",
    subtitle = "Observed x = 8 out of n = 170",
    x = expression(pi),
    y = "Density"
  )
```

```{r}
#Posterior Summary calculation
qbeta(0.5, shape1 = 36.40 + 8, shape2 = 38.58 + 170 - 8)


qbeta(p = c(0.025, 0.975), shape1 = 36.40 + 8, shape2 = 38.58 + 170 - 8)

hdi(qbeta, credMass = 0.95, shape1 = 36.40 + 8, shape2 = 38.58 + 170 - 8)
```

```{r}
#Bayesian P value
pbeta(0.7/1.7, shape1 = 44.40, shape2 = 200.58, lower.tail = FALSE)
```

```{r}
#Bootstrap CI
set.seed(342)

n <- 170
v <- 8
pi_hat <- v/n
psi_hat <- (n - 2*v)/(n - v)

B <- 10000

t <- rbinom(n = B, size = n, prob = pi_hat)

psi <- (n - 2*t) / (n - t)

c(lower = quantile(psi, 0.025),upper = quantile(psi, 0.975))
```

```{r}
#Fisher Information CI
n <- 170
v <- 8
psi_hat <- (n - 2*v) / (n - v)

I <- 8/(1 - psi_hat)^2 + 170/(2 - psi_hat)^2

SE <- 1/sqrt(I)

lower <- psi_hat - 1.96 * SE
upper <- psi_hat + 1.96 * SE
c(lower,upper)
```

# Proofs and Working

To show that T is a binomial random variable, we have X = number of Covid cases in the vaccine group and Y = number in placebo group. Since n is large for both we have,
$$X \sim Pois(n_1, \pi_v) $$
$$Y \sim Pois(n_2, \pi_p) $$

The PMF of T can be represented as,

$$P(T = t) = P(X = t | X + Y = n)$$

$$= \frac{P(X = t) \times P(Y = n-t)}{P(X+Y = n)}$$

$$= \frac{\frac{e^{-n_1\pi_v}(n_1\pi_v)^t}{t!} \times \frac{e^{-n_2\pi_p}(n_2\pi_p)^{n-t}}{(n-t)!}}{\frac{e^{-(n_1\pi_v + n_2\pi_p)}(n_1\pi_v + n_2\pi_p)^n}{n!}}$$

$$= \frac{n!}{t!(n-t)!} \times \frac{e^{-n_1\pi_v}(n_1\pi_v)^t \times e^{-n_2\pi_p}(n_2\pi_p)^{n-t}}{e^{-(n_1\pi_v + n_2\pi_p)}(n_1\pi_v + n_2\pi_p)^n}$$

$$= \binom{n}{t} \times \frac{e^{-(n_1\pi_v + n_2\pi_p)}(n_1\pi_v)^t(n_2\pi_p)^{n-t}}{e^{-(n_1\pi_v + n_2\pi_p)}(n_1\pi_v + n_2\pi_p)^n}$$

$$= \binom{n}{t} \times \frac{(n_1\pi_v)^t(n_2\pi_p)^{n-t}}{(n_1\pi_v + n_2\pi_p)^n}$$

$$= \binom{n}{t} \times \frac{(n_1\pi_v)^t(n_2\pi_p)^{n-t}}{(n_1\pi_v + n_2\pi_p)^{t+n-t}}$$

$$= \binom{n}{t} \times \frac{(n_1\pi_v)^t(n_2\pi_p)^{n-t}}{(n_1\pi_v + n_2\pi_p)^t(n_1\pi_v + n_2\pi_p)^{n-t}}$$

$$= \binom{n}{t} \left(\frac{n_1\pi_v}{n_1\pi_v + n_2\pi_p}\right)^t \left(\frac{n_2\pi_p}{n_1\pi_v + n_2\pi_p}\right)^{n-t}$$

$$= \binom{n}{t} \pi^t (1-\pi)^{n-t}$$

To simply our estimator $\pi$ we have by definition of conditional probability,
$$\pi = P(Vaccine | Covid19) =\frac{ P(Vaccine) \land P(Covid19)}{P(Covid19)}$$
$$=\frac{ P(Vaccine)  P(Covid19 | Vaccine)}{P(Vaccine)P(Covid19 | vaccine) + P(placebo)P(Covid19 | placebo)}$$

$$ = \frac{\frac{n1}{n1 + n2}\pi}{\frac{n_1 \pi_v + n_2 \pi_p}{n_1 + n_2}}$$
$$ = \frac{n_1 \pi_v}{n_1 \pi_v + n_2 \pi_p}$$

To represent $\psi$ in terms of $\pi$, 
$$\psi = \frac{1 - 2\pi}{1 - \pi}$$
$$(1 - \pi)\psi = 1 - 2\pi$$
$$2\pi- \pi\psi = 1 - \psi$$
$$\pi(2 - \psi) = 1 - \psi$$
$$\pi = \frac{1 - \psi}{2 - \psi}$$

To calculate the MLE we do the following,
$$L(\pi) = \binom{n}{k}\pi^k(1-\pi)^{n - k} =\binom{170}{8}\pi^8(1-\pi)^{162} $$
$$L^*(\psi)  =\binom{170}{8}(\frac{1 - \psi}{2 - \psi})^8(1-\frac{1 - \psi}{2 - \psi})^{162} $$
$$l^*(\psi)  =log(\binom{170}{8}) + 8log(1 - \psi) - 8log(2 - \psi) +162log(1-\frac{1 - \psi}{2 - \psi}) $$
$$ =log(\binom{170}{8}) + 8log(1 - \psi) - 8log(2 - \psi) +162log(\frac{1 }{2 - \psi}) $$
$$ =log(\binom{170}{8}) + 8log(1 - \psi) - 8log(2 - \psi) +162log(1) - 162log(2 - \psi) $$
$$ =log(\binom{170}{8}) + 8log(1 - \psi) +162log(1) - 170log(2 - \psi) $$
$$l^{*'}(\psi)  = -\frac{8}{1 - \psi} + \frac{170}{2 - \psi} $$
$$-\frac{8}{1 - \psi} + \frac{170}{2 - \psi} = 0$$
$$\frac{170}{2 - \psi} = \frac{8}{1 - \psi}$$
$$170 - 170\psi = 16 - 8\psi$$
$$154 = 162\psi$$
$$\hat \psi^{mle} = \frac{154}{162} = 0.95061$$

We will now calculate the interval using this distribution of our estimate,
The fisher expected fisher information is given by,
$$I(\psi_0) = E[-\frac{d^2}{d\psi^2}l^*(\psi)]_{\psi = \psi_0}$$
$$-\frac{d^2}{d\psi^2}l^*(\psi) = \frac{8}{(1 - \psi)^2} + \frac{170}{(2 - \psi)^2}$$
$$E[-\frac{d^2}{d\psi^2}l^*(\psi)]_{\psi = \psi_0} = \frac{8}{(1 - \psi)^2} + \frac{170}{(2 - \psi)^2}$$
$$ = \frac{8}{(1 - \frac{154}{162})^2} + \frac{170}{(2 - \frac{154}{162})^2}$$
$$ = \frac{583929}{170}$$
We can calculate a 95% confidence interval with,
$$\hat \psi^{mle} \pm z_{\alpha/2}\sqrt{\frac{1}{I(\hat \psi^{mle})}}$$

So we get the 95% confidence interval using R as,
(0.9171746, 0.9840599)